\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-14 Manuscript/manuscript.tex
Made some more small changes to tables

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{The top graph shows the reaction times of athletes who competed at 
	least once each at the 2019 and 2022 World Championships. The middle graph 
	shows the reaction times of athletes who competed at least once each at the 
	2022 and 2023 World Championships. The bottom graph shows the reaction times 
	of athletes who competed at least once at their country's
  2022 national championships and the 2022 World Championships.}
  \label{fig:RankScatterplots}
\end{figure}
\of{Need to go through and edit this}

Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
with the rank-based methods (described in Section~\ref{sec:rank} with results in 
Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
times for athletes who competed at the 2022 World Championships
are compared to their reaction times at another at which they competed. 
It is worth noting that each instance of the 2022 data is different, because
those who competed at both the 2019 and 2022 World Championships will be
different athletes than those who competed at both the 2022 and 2023 World
Championships. This data was only for athletes who competed
in at least one race at each championship, however it is possible for athletes
to have competed in one race in 2019 and two races in 2022. The 2022 World
Championship heats and semifinals both had on average lower reaction times than
their respective other (2019, 2023 or 2022 national) races, and the 2022 World 
Championship finals were substantially faster.  It
is also worth noting is that the fastest reaction time in the 2022 World
Championships Finals and Semifinals both came from Devon Allen, but the 
difference between 0.101 and 0.099
seconds was the difference in what caused him to be disqualified.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
  \label{fig:Boxplot}
\end{figure}
\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three models fitted to the
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
    \midrule
    Excluding 2022 &
      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.868$ & 0.136\\[1ex]
    Including 2022 &
    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.149\\
    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.881$ & 0.134\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than threshold 0.08,
  0.09, and 0.10 seconds based on the two effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\ 
   \midrule
   Excluding 2022 & $3.31\cdot10^{-5}$ & $4.46\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
   Including 2022 & $4.16\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

The GLMM with both venue and heat random effects helps us to assess how extreme
a reaction time below a given threshold is. The probability of observing a reaction
time below a threshold, assuming no intentional false starts, can be
approximated by generating a large number of realizations from the fitted
GLMM. We used 10 millions realizations in the analysis. 
Table~\ref{tab:Sim_probability} summarizes the estimated probabilities of
observing a reaction time below 0.08, 0.09, and 0.10 seconds under two different
scenarios: one excluding and the other including data from 2022. Apparently,
excluding 2022 lowered the probability of observing a fast reaction time, but
the difference is not big in magnitude. For example, the probability for
seeing a reaction time below 0.10 seconds changes from $3.54\cdot 10^{-3}$ to
$4.20\cdot 10^{-3}$ when 2022 is included. These values correspond to roughly one in
every 230 starts and 283 starts, respectively. The table shows that changing the
reaction time barrier from 0.1 seconds to 0.08 seconds drastically reduces the
chance of observing a reaction time below the barrier, from one in roughly every
283 starts to one in every 29851 starts using data with 2022 included. The
results substantiates the recommendations by \citet{komi2009iaaf}.

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on tail probabilities.}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\ 
   \midrule
   Excluding 2022 & $0.094$ & $0.084$ \\ 
   Including 2022 & $0.092$ & $0.083$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}

>>>>

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{This collection of scatterplots of data used in the rank comparison
  section shows that times in 2022 were on average lower. On the horizontal
  axis below each graph "H", "S", and "F" refer to the heats, semifinals, and 
  finals respectively. Please note that in the last row the 2022 times are to
  the right of the 2023 times.}
  \label{fig:RankScatterplots}
\end{figure}

Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
with the rank-based methods (described in Section~\ref{sec:rank} with results in 
Section~\ref{subsec:Results_Rank}). In each of the six graphs, the reaction 
times for athletes who competed at the 2022 World Championships
are compared to their reaction times at another at which they competed. 
It is worth noting that each instance of the 2022 data is different, because
those who competed at both the 2019 and 2022 World Championships will be
different athletes than those who competed at both the 2022 and 2023 World
Championships. This data was only for athletes who competed
in at least one race at each championship, however it is possible for athletes
to have competed in one race in 2019 and two races in 2022. The 2022 World
Championship heats and semifinals both had on average lower reaction times than
their respective other (2019, 2023 or 2022 national) races, and the 2022 World 
Championship finals were substantially faster.  It
is also worth noting is that the fastest reaction time in the 2022 World
Championships Finals and Semifinals both came from Devon Allen, but the 
difference between 0.101 and 0.099
seconds was the difference in what caused him to be disqualified.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle
  and 100 meter dash.}
  \label{fig:Boxplot}
\end{figure}
\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three models fitted to the
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
    \midrule
    Excluding 2022 &
      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.868$ & 0.136\\[1ex]
    Including 2022 &
    Venue Effect Only & $-3798.8$ &   & 0.054 & $-1.871$ & 0.149\\
    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.881$ & 0.134\\
    \bottomrule
  \end{tabular}
\end{table}
\of{needs to be updated}

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than threshold 0.08,
  0.09, and 0.10 seconds based on the two effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\ 
   \midrule
   Excluding 2022 & $1.10\cdot10^{-5}$ & $2.34\cdot10^{-4}$ &  $2.36\cdot10^{-3}$  \\ 
   Including 2022 & $1.71\cdot10^{-5}$ & $3.01\cdot10^{-4}$ & $2.90\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}
\of{table has been updated}

The GLMM with both venue and heat random effects helps us to assess how extreme
a reaction time below a given threshold is. The probability of observing a reaction
time below a threshold, assuming no intentional false starts, can be
approximated by generating a large number of realizations from the fitted
GLMM. We used 10 millions realizations in the analysis. 
Table~\ref{tab:Sim_probability} summarizes the estimated probabilities of
observing a reaction time below 0.08, 0.09, and 0.10 seconds under two different
scenarios: one excluding and the other including data from 2022. Apparently,
excluding 2022 lowered the probability of observing a fast reaction time, but
the difference is not big in magnitude. For example, the probability for
seeing a reaction time below 0.10 seconds changes from $2.36\cdot 10^{-3}$ to
$2.90\cdot 10^{-3}$ when 2022 is included. These values correspond to one in
every 344 starts and 423 starts, respectively. The table shows that changing the
reaction time barrier from 0.1 seconds to 0.08 seconds drastically reduces the
chance of observing a reaction time below the barrier, from one in roughly every
283 starts to one in every 58480 starts using data with 2022 included. The
results substantiates the recommendations by \citet{komi2009iaaf}.

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on tail probabilities.}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\ 
   \midrule
   Excluding 2022 & $0.096$ & $0.087$ \\ 
   Including 2022 & $0.095$ & $0.086$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}
\of{table has been updated}
